---
title: Pokémon Statistics
author: Aaron Lang
output:
  html_notebook:
    code_folding: hide
    toc: yes
    toc_float: true
  html_document:
    code_folding: hide
    df_print: kable
    toc: yes
    toc_float: true
---

- Descriptive statistics involves analysing and plotting data
- Inferential statistics allows us to infer information about some population from a sample

- μ : Population mean
- σ : Population standard deviation
- x- : Sample mean
- s : Sample s.d.

- Analysing height is 1D
- Analysing height against type is 2D

### Variable Types

- Continuos numeric
  - Height
  - Weight
- Discrete numeric
  - Number of species with type
  - Number of species in family
- Nominal categorical
  - Type
  - Attack > Defense
- Ordinal categorical
  - Evolution level
  - Evolution level relative to family mean
  - Generation

Indentifying the var type informs analysis and graphing:

    Linear regression if the outcome is continuous numeric

    Logistic regression if the outcome is dichotomous

    Multinomial regression if the outcome is nominal categorical

    Proportional odds logistic regression if the outcome is ordinal categorical

    Poisson or negative binomial regression if the outcome is a count or frequency

    Cox proportional hazards model (or accelerated failure time models) if the outcome is a censored time-to-event e.g. survival time.

Skewed distributions are poorly described by mean and sd, better by median and IQR


```{r include=FALSE}
library(dplyr)
library(ggplot2)
library(moments)
pokemon <- read.csv(file = "pokemon.csv")
pokemon$is_legendary <- as.logical(pokemon$is_legendary)
```
## Categorical Data

Let's look at legends and generation

First, generation:

```{r}
pokemon <- pokemon %>% mutate(type_count = ifelse(type2 == '', 1, 2))
# pokemon %>% head %>% select(name, type1, type2, type_count)
table(pokemon$generation)
ggplot(pokemon, aes(generation)) + geom_bar()
barplot(table(pokemon$generation))
```
Now legends
```{r}
pokemon <- pokemon %>% mutate(gender_is_na = is.na(percentage_male))
table_is_legendary = table(pokemon$is_legendary)
table_is_legendary
prop.table(table_is_legendary)
```
And both together
```{r}
table_2way = table(pokemon$generation, pokemon$is_legendary)
table_2way
```
This doesn't make it very easy to compare across generations, so we can try this
```{r}
tmp = table_2way * 100
tmp = apply(tmp, 2, function(x) sprintf('%.1f%%', x))
dimnames(tmp) =  list(1:7, c(FALSE, TRUE))
tmp
```
```{r}
mosaicplot(table_2way, dir=c("v","h"))
barplot(table(pokemon$is_legendary, pokemon$generation))
```
## Continous Data
```{r}
hist(pokemon$hp)
plot(density(pokemon$hp))
hist(pokemon$percentage_male)
#plot(density(pokemon$percentage_male))
```
## Continuos x Categorical data (Bivariate)

```{r}
boxplot(data = pokemon, speed ~ generation)
boxplot(data = pokemon, weight_kg ~ generation)
```

## Continuos * Continuos

```{r}
plot(pokemon$height_m, pokemon$weight_kg)
plot(pokemon$attack, pokemon$defense)
gen1 = pokemon %>% filter(generation == 1)
plot(gen1$height_m, gen1$weight_kg)
plot(gen1$attack, gen1$defense)
pairs(data=gen1, ~attack+defense+speed+hp)
library(car)
scatterplotMatrix(~attack+defense+speed+hp, data=gen1)
```
Covariance is the means of the product of the differences
```{r}
mean_attack = mean(gen1$attack)
sd_attack = sd(gen1$attack)
mean_defense = mean(gen1$defense)
sd_defense = sd(gen1$defense)
gen1$attack_sub_mean = gen1$attack - mean_attack
gen1$defense_sub_mean = gen1$defense - mean_defense
products = gen1$attack_sub_mean * gen1$defense_sub_mean
covn = mean(products)
covn1 = sum(products)/(nrow(gen1) - 1)
covn # 401.6
covn1 # 404.29
cov(gen1$attack, gen1$defense) # 404.29
covn1/(sd_attack*sd_defense) # pearson 0.464
cor(gen1$attack, gen1$defense) # pearson 0.464
```
# Height

```{r}
summary(pokemon$height_m)
sd(pokemon$height_m, na.rm = TRUE)
skewness(pokemon$height_m, na.rm = TRUE)
kurtosis(pokemon$height_m, na.rm = TRUE)
```
```{r}
pokemon %>% arrange(height_m) %>% head %>% select(generation, pokedex_number, name, height_m)
pokemon %>% arrange(-height_m) %>% head %>% select(generation, pokedex_number, name, height_m)
```
```{r}
ggplot(pokemon, aes(height_m)) + geom_density() + scale_x_log10()
```


# Weight

```{r}
summary(pokemon$weight_kg)
```
```{r}
pokemon %>% arrange(weight_kg) %>% head %>% select(generation, pokedex_number, name, weight_kg)
pokemon %>% arrange(-weight_kg) %>% head %>% select(generation, pokedex_number, name, weight_kg)
```
```{r}
ggplot(pokemon, aes(weight_kg)) + geom_density() + scale_x_log10()
```

# Density

```{r}
pokemon <- pokemon %>% mutate(density = weight_kg/height_m)
summary(pokemon$density)
```
```{r}
pokemon %>% arrange(density) %>% head %>% select(generation, pokedex_number, name, density)
pokemon %>% arrange(-density) %>% head %>% select(generation, pokedex_number, name, density)
```
```{r}
ggplot(pokemon, aes(density)) + geom_density() + scale_x_log10()
```

# Body Mass Index (BMI)

```{r}
pokemon <- pokemon %>% mutate(bmi = weight_kg/(height_m^2))
summary(pokemon$bmi)
```
```{r}
pokemon %>% arrange(bmi) %>% head %>% select(generation, pokedex_number, name, bmi)
pokemon %>% arrange(-bmi) %>% head %>% select(generation, pokedex_number, name, bmi)
```
```{r}
ggplot(pokemon, aes(bmi)) + geom_density() + scale_x_log10()
```